---
title: "CoGAPS"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
        pandoc_args: [
         "+RTS", "-k16g",
         "-RTS"
         ]
        
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 20px;
  color: DarkBlue;
}
h2 { /* Header 2 */
  font-size: 18px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 16px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Working directory:
```{r,message=FALSE,warning=FALSE}
print(getwd())
```

# Check required packages
```{r,message=FALSE,warning=FALSE}
# function for installing needed packages
installpkg <- function(x){
    r = getOption("repos")
    r["CRAN"] = "http://cran.us.r-project.org"
    options(repos = r)
    if(x %in% rownames(installed.packages())==FALSE) {
        if(x %in% rownames(available.packages())==FALSE) {
            if (!requireNamespace("BiocManager", quietly = TRUE))
               install.packages("BiocManager")
               BiocManager::install(x,ask = FALSE)
        } else {
            install.packages(x)
        }
        paste(x, "package is installed...")
    } else {
        paste(x,"package already installed...")
    }
}

# install necessary packages
required_packages  <- c('ggplot2','Seurat','SeuratWrappers','plotly','CoGAPS','stringr')
lapply(required_packages,installpkg)
```

# Using Seurat wrapped CoGAPS to identify seurat generatd cell clusters, display first page of the figures
```{r,message=FALSE,warning=FALSE,results='hide'}
library(Seurat)
library(SeuratWrappers)
library(CoGAPS)
library(ggplot2)
load("../seurat/seurat.Rdata")
n <- length(levels(seurat$seurat_clusters))
params <- CogapsParams(singleCell = TRUE, sparseOptimization = TRUE, seed = 66, nIterations = 5000, 
    nPatterns = n+1, distributed = "genome-wide")
params <- setDistributedParams(params, nSets = 3)
cog <-RunCoGAPS(seurat, params = params)
write.csv(cog@reductions$CoGAPS[],"cogaps_cluster_featureloadings.csv",quote=F)
write.csv(cog@reductions$CoGAPS[[]],"cogaps_cluster_samplefactors.csv",quote=F)
pdf("cogaps_cluster.pdf")
p <- length(colnames(cog@reductions$CoGAPS))
for (i in 1:p)
{
a <- paste("CoGAPS_",i,sep="")
b<- VlnPlot(cog,features=a)
plot(b)
}
dev.off()
```
<button class="btn btn-primary" data-toggle="collapse" data-target="#Spectra">Open cogaps_cluster.pdf</button> 
<div id="Spectra" class="collapse">    
<object data="cogaps_cluster.pdf" width="750px" height="750px">
    <embed src="cogaps_cluster.pdf">
    </embed>
</object>
</div>


# Using Seurat wrapped CoGAPS to identify SingleR generatd cell types
```{r,message=FALSE,warning=FALSE,results='hide'}
ano <- read.csv("../singler/singleR.csv",header=T,row.names=1)
library(stringr) 
rownames(ano)<-str_replace_all(rownames(ano),"\\.","-")
ano <- ano[order(rownames(ano)),]
seurat@meta.data$singleR.celltype <- ano$labels
n <- length(unique(seurat@meta.data$singleR.celltype))
params2 <- CogapsParams(singleCell = TRUE, sparseOptimization = TRUE, seed = 66, nIterations = 5000, 
    nPatterns = n+1, distributed = "genome-wide")
params2 <- setDistributedParams(params2, nSets = 3)
seurat2 <-seurat
Idents(seurat2) <- "singleR.celltype"
cog2 <-RunCoGAPS(seurat2, params = params2)
write.csv(cog2@reductions$CoGAPS[],"cogaps__celltype_featureloadings.csv",quote=F)
write.csv(cog2@reductions$CoGAPS[[]],"cogaps_celltype_samplefactors.csv",quote=F)
pdf("cogaps_celltype.pdf")
p <- length(colnames(cog2@reductions$CoGAPS))
for (i in 1:p)
{
a <- paste("CoGAPS_",i,sep="")
b <- VlnPlot(cog2,features=a) + theme(legend.position = 'none')
plot(b)
}
dev.off()

```
<button class="btn btn-primary" data-toggle="collapse" data-target="#Spectra2">Open cogaps_celltype.pdf</button> 
<div id="Spectra2" class="collapse">    
<object data="cogaps_celltype.pdf" width="750px" height="750px">
    <embed src="cogaps_celltype.pdf">
    </embed>
</object>
</div>

```{r,message=FALSE,warning=FALSE}
save.image("cogaps.Rdata")
```
