---
title: "combining of seurat, singleR and cogaps resuts"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
        pandoc_args: [
         "+RTS", "-k16g",
         "-RTS"
         ]

---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 20px;
  color: DarkBlue;
}
h2 { /* Header 2 */
  font-size: 18px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 16px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Working directory:
```{r,message=FALSE,warning=FALSE}
print(getwd())
```

# Check required packages
```{r,message=FALSE,warning=FALSE}
# function for installing needed packages
installpkg <- function(x){
    r = getOption("repos")
    r["CRAN"] = "http://cran.us.r-project.org"
    options(repos = r)
    if(x %in% rownames(installed.packages())==FALSE) {
        if(x %in% rownames(available.packages())==FALSE) {
            if (!requireNamespace("BiocManager", quietly = TRUE))
               install.packages("BiocManager")
               BiocManager::install(x,ask = FALSE)
        } else {
            install.packages(x)
        }
        paste(x, "package is installed...")
    } else {
        paste(x,"package already installed...")
    }
}

# install necessary packages
required_packages  <- c('ggplot2','Seurat','rmarkdown','stringr')
lapply(required_packages,installpkg)
```

# Read in Seurat data
```{r,message=FALSE,warning=FALSE}
library(Seurat)
load("../seurat/seurat.Rdata")
cl <- read.csv("../seurat/cluster.csv",header=F,row.names=1)
cl <- cl[order(rownames(cl)),,drop=F]
```

# Read in SingleR cell annotation infomration
```{r,message=FALSE,warning=FALSE}
ano <- read.csv("../singler/singleR.csv",header=T,row.names=1)
library(stringr) 
rownames(ano)<-str_replace_all(rownames(ano),"\\.","-")
ano <- ano[order(rownames(ano)),]
seurat@meta.data$singleR.celltype <- ano$labels
```

# Integrate the cluster and  cell annotation infomration
```{r,message=FALSE,warning=FALSE}
group <- cbind(cl,ano$labels)
colnames(group) <- c("cluster","estimated.cell.type")
write.csv(group,"cluster_annotation.csv",quote = F)
library(rmarkdown)
paged_table(group)
```

# Plots annotated with clusters from Seurat

## UMAP plot annotated with clusters from Seurat
```{r,message=FALSE,warning=FALSE}
library(ggplot2)
DimPlot(seurat,reduction="umap")
```
## TSNE plot annotated with clusters from Seurat
```{r,message=FALSE,warning=FALSE}
DimPlot(seurat,reduction="tsne")
```

<button class="btn btn-primary" data-toggle="collapse" data-target="#Spectra">Cogaps patterns annotated with clusters from Seurat</button> 
<div id="Spectra" class="collapse">    
<object data="cogaps_cluster.pdf" width="750px" height="750px">
    <embed src="cogaps_cluster.pdf">
    </embed>
</object>
</div>

# Plots annotated with cell types from SingleR

## UMAP plot annotated with cell types from SingleR
```{r,message=FALSE,warning=FALSE}
seurat2 <-seurat
Idents(seurat2) <- "singleR.celltype"
DimPlot(seurat2,reduction="umap") + theme(legend.position="bottom",legend.text=element_text(size=6)) + guides(colour = guide_legend(override.aes = list(size=2)))
```
## TSNE plot annotated with cell types from SingleR
```{r,message=FALSE,warning=FALSE}
DimPlot(seurat2,reduction="tsne") + theme(legend.position="bottom",legend.text=element_text(size=6)) + guides(colour = guide_legend(override.aes = list(size=2)))
```
<button class="btn btn-primary" data-toggle="collapse" data-target="#Spectra2">Cogaps patterns annotated with cell types from SingleR</button> 
<div id="Spectra2" class="collapse">    
<object data="cogaps_celltype.pdf" width="750px" height="750px">
    <embed src="cogaps_celltype.pdf">
    </embed>
</object>
</div>

```{r,message=FALSE,warning=FALSE}
save.image("combine.Rdata")
```
