---
title: "diem"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
        
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 20px;
  color: DarkBlue;
}
h2 { /* Header 2 */
  font-size: 18px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 16px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Use DIEM to clean up cellranger snRNA-seq raw data and compare with the cellranger filtered data

# Working directory:
```{r,message=FALSE,warning=FALSE}
print(getwd())
```

# Check required packages
```{r,message=FALSE,warning=FALSE}
# function for installing needed packages
installpkg <- function(x){
    r = getOption("repos")
    r["CRAN"] = "http://cran.us.r-project.org"
    options(repos = r)
    if(x %in% rownames(installed.packages())==FALSE) {
        if(x %in% rownames(available.packages())==FALSE) {
            if (!requireNamespace("BiocManager", quietly = TRUE))
               install.packages("BiocManager")
               BiocManager::install(x,ask = FALSE)
        } else {
            install.packages(x)
        }
        paste(x, "package is installed...")
    } else {
        paste(x,"package already installed...")
    }
}

# install necessary packages
required_packages  <- c('ggplot2','diem','Seurat')
lapply(required_packages,installpkg)
```

# Read in filtered count data
```{r,message=FALSE,warning=FALSE}
library(diem)
library(Seurat)
counts_f <- read_10x("filtered_feature_bc_matrix")
sce_f <- create_SCE(counts_f)
ncol(counts_f)
nrow(counts_f)
```

# Read in raw count data
```{r,message=FALSE,warning=FALSE}
counts <- read_10x("raw_feature_bc_matrix")
sce <- create_SCE(counts)
ncol(counts)
nrow(counts)
```

# Add MT% (mitochondria) and MALAT1% (nuclear-localized lincRNA)
```{r,message=FALSE,warning=FALSE}
mt_genes <- grep(pattern="^mt-", x=rownames(sce@gene_data), ignore.case=TRUE, value=TRUE)
sce <- get_gene_pct(x = sce, genes=mt_genes, name="pct.mt")
malat <- grep(pattern="^malat1$", x=rownames(sce@gene_data), ignore.case=TRUE, value=TRUE)
sce <- get_gene_pct(x = sce, genes=malat, name="MALAT1")
```

# Plot total counts ranked
```{r,message=FALSE,warning=FALSE}
library(ggplot2)
#cellranger filtered counts
barcode_rank_plot(sce_f)
#cellranger raw counts
barcode_rank_plot(sce)
```

# DIEM steps
```{r,message=FALSE,warning=FALSE}
sce <- set_debris_test_set(sce)
sce <- filter_genes(sce)
sce <- get_pcs(sce)
sce <- init(sce)
sce <- run_em(sce)
sce <- assign_clusters(sce)
sce <- estimate_dbr_score(sce)
```

# Call targets using debris score for single-nucleus data
```{r,message=FALSE,warning=FALSE}
sce <- call_targets(sce, thresh_score = 0.5)
seur <- convert_to_seurat(sce)
seur
```

# Get final cleaned cell barcodes for cellranger reanalysis
```{r,message=FALSE,warning=FALSE}
counts2 <- counts_f[,colnames(counts_f)%in%colnames(seur)]
ncol(counts2)
barcode <- colnames(counts2)
barcode <- as.data.frame(barcode)
colnames(barcode) <- "Barcode"
barcode $ Barcode <- paste(barcode $ Barcode,"-1",sep="")
write.csv(barcode,"barcode.csv",row.names=F,quote = F)
                    
save.image("diem.Rdata")
```


