---
title: "singleR"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
        
---

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 20px;
  color: DarkBlue;
}
h2 { /* Header 2 */
  font-size: 18px;
  color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 16px;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Working directory:
```{r,message=FALSE,warning=FALSE}
print(getwd())
```

# Check required packages
```{r,message=FALSE,warning=FALSE}
# function for installing needed packages
installpkg <- function(x){
    r = getOption("repos")
    r["CRAN"] = "http://cran.us.r-project.org"
    options(repos = r)
    if(x %in% rownames(installed.packages())==FALSE) {
        if(x %in% rownames(available.packages())==FALSE) {
            if (!requireNamespace("BiocManager", quietly = TRUE))
               install.packages("BiocManager")
               BiocManager::install(x,ask = FALSE)
        } else {
            install.packages(x)
        }
        paste(x, "package is installed...")
    } else {
        paste(x,"package already installed...")
    }
}

# install necessary packages
required_packages  <- c('ggplot2','data.table','ExperimentHub','SingleR','rmarkdown','celldex')
lapply(required_packages,installpkg)
```

# Read in Seurat log normalized (not scaled) counts data of variable genes
```{r,message=FALSE,warning=FALSE}
library(data.table)
system.time(data <- (fread("../seurat/log_normalized_data_var_genes.csv",header=T)))
data <- as.data.frame(data)
rownames(data) <- data $V1
data <- data[,2:ncol(data)]
data <- as.matrix(data)
```

# Harmonizing labels of references: probabilities close to 1 indicate there is a 1:1 relation between that pair of labels; on the other hand, an all-zero probability vector indicates that a label is unique to a particular reference.
```{r,message=FALSE,warning=FALSE}
library(ExperimentHub)
setExperimentHubOption("PROXY", Sys.getenv("http_proxy"))
library(SingleR)
library(celldex)
hpca.se <- celldex::HumanPrimaryCellAtlasData()
hpca.se
bp.se <- BlueprintEncodeData()
bp.se
matched <- matchReferences(bp.se, hpca.se,
    bp.se$label.main, hpca.se$label.main)
pheatmap::pheatmap(matched, col=viridis::plasma(100))
```

# Annotate data using singleR with multiple built-in references by comparing scores across the union of markers
```{r,message=FALSE,warning=FALSE}
pred.combined <- SingleR(test = data, ref = list(BP=bp.se, HPCA=hpca.se), labels = list(bp.se$label.main,hpca.se$label.main),de.method="wilcox")
library(rmarkdown)
paged_table(as.data.frame(pred.combined))
write.csv(pred.combined,"singleR.csv",quote = F) 
```

# Annotation visualization. Low-quality assignments are replaced with NA.
```{r,message=FALSE,warning=FALSE}
plotScoreHeatmap(pred.combined,scores.use = 0,calls.use=0)
plotScoreHeatmap(pred.combined,scores.use = 1,calls.use=1)
plotScoreHeatmap(pred.combined,scores.use = 2,calls.use=2)

save.image("singler.Rdata")
```
